Hidden Multiple Comparisons Increase Forensic Error Rates

### 2024-05-15 PNAS/index.Rnw
Trying to ensure the PNAS version works the first time... lol
Movinng all the pnas-required files into the damn folder so that everything works... hopefully?
Final updates, hopefully.

\title{Hidden Multiple Comparisons Increase Forensic Error Rates}

@@

\section*{Calculating the Number of Comparisons}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{../fig/wire-blade-comparison-label.png}
\includegraphics[width=\columnwidth]{../fig/wire_cutter_wires.png}
\caption{(Top) A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match). Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope. (Bottom) RJ45 Crimp tool with a $\Sexpr{b/10}$ cm razor blade used for cutting. 1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}\label{fig:wire-blade}
\end{figure}

@@

\section*{Probability of False Discoveries}
There are at least two components of the false discovery rate (FDR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective rules for how much similarity is sufficient for an identification.
Assuming that lab procedure errors are not a factor in studies, we use reported error rates from three open-set studies of striated evidence\citep{bajic2020, mattijssen2021,best2022} to obtain a ballpark estimate of the coincidental match rate of a single wire-cut comparison.
These studies have FDRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies weighted by sample size yields an FDR of \Sexpr{sprintf("%.2f", fpr)}.
For a single-comparison FDR of $e$, the family-wise FDR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
\Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates. With an error rate of 0.07, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false discovery error.
To conduct a search of a modestly sized database with 1000 entries, the initial FDR cannot
exceed 1 in 10,000 to guarantee a family-wise total false discovery error of at most 10\%.

>>>>

\graphicspath{{../fig/}}
\title{Hidden Multiple Comparisons Increase Forensic Error Rates}

@@

\section*{Methods: Calculating the Number of Comparisons}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{full-wire-cutter-fig.pdf}
\caption{(Top) A comparison between a wire and a blade cut requires sliding the wire along the entire blade cut length to determine the best match (or whether there is a match). Surfaces shown are rendered 2D topographical scans of a wire and blade cut taken with a confocal light microscope. (Bottom) RJ45 Crimp tool with a $\Sexpr{b/10}$ cm razor blade used for cutting. 1 mm and 2 mm diameter aluminum wires cut with the pliers are shown in a box in the top right corner.}\label{fig:tool-pic}\label{fig:wire-blade}
\end{figure}

@@

\section*{Probability of False Discoveries}
There are at least two components of the false discovery rate (FDR): identifying two pieces of evidence that have similar characteristics but are from different sources (a coincidental match) and procedural failures (e.g.\ lab process errors) \citep[p 50]{pcast}.
In objective disciplines with standardized evaluation rules (e.g.\ DNA), these sources can be distinguished.
However, in toolmark examination, no objective evaluation rules are used; examiners testify based on subjective rules for how much similarity is sufficient for an identification.
Assuming that lab procedure errors are not a factor in studies, we use reported error rates from three open-set studies of striated evidence\citep{bajic2020, mattijssen2021,best2022} to obtain a ballpark estimate of the coincidental match rate of a single wire-cut comparison.
These studies have FDRs between 0.0045 \citep{best2022} and 0.072 \citep{mattijssen2021}; pooling data from these studies weighted by sample size yields an FDR of \Sexpr{sprintf("%.2f", fpr)}.
For a single-comparison FDR of $e$, the family-wise FDR for $n$ comparisons, $E_n$ is $1 - [1-e]^n$.
\Cref{tab:family-wise} shows the impact the number of comparisons has on these published error rates. With an error rate of 0.007, as suggested by Bajic (2020), examiners can make up to 14 comparisons, i.e.\ even the simple example in this paper exceeds an upper bound of 10\% for the family wise false discovery error.
To conduct a search of a modestly sized database with 1000 entries, the initial FDR cannot
exceed 1 in 10,000 to guarantee a family-wise total false discovery error of at most 10\%.
